---
title: "Final_Project"
author: "Shahin Shahsavari"
date: "4/15/2022"
output: html_document
---

# 2017 Medicare PartD Claims

```{r}
library(tidyverse)
```

1. Load the `~/Day4/datasets/medicare_claims.tsv` file in R

```{r}
medicare_claims <- read_tsv("~/Day4/datasets/medicare_claims.tsv")
```

2. Identify the columns of this data frame using glimpse()

```{r}
glimpse(medicare_claims)
```

3. Use `count()` on the column that identifies the state
  * the count() function counts the number of appearances of each element in a column

```{r}
medicare_claims %>%
  count(provider_state)
```

4. Filter the dataframe and find all the BUPRENORPHINE claims (hint: This is in the drug_name column).

```{r}
medicare_claims %>%
  filter(drug_name == "BUPRENORPHINE")
```

  
5. Which State has the highest number of BUPRENORPHINE claims?
  * you can use arrange() to sort your column

```{r}
medicare_claims %>%
  filter(drug_name == "BUPRENORPHINE") %>%
  count(provider_state) %>%
  arrange(desc(n))
```


6. We just found out that sometimes the providers use the terms "BUPRENORPHINE HCL" or "BUPRENORPHINE-NALOXONE". Revise your filter to include those as well.

```{r}
buprenorphine_claims <- medicare_claims %>%
  filter(drug_name == "BUPRENORPHINE" | drug_name == "BUPRENORPHINE HCL" | drug_name == "BUPRENORPHINE-NALOXONE")
```


7. Use `geom_bar()` to find the aggregate count of buprenorphine claims with the state names as your x axis.
       * generate the count table first, then use geom_bar to plot the aggregate counts

```{r}
buprenorphine_state_count <- buprenorphine_claims %>%
  count(provider_state)
ggplot() +
  geom_bar(data = buprenorphine_state_count,
           aes(x = reorder(provider_state,-n), y = n),
           stat = "identity")
```


Next, we would like to find out which providers (by specialty) prescribe buprenorphine the most.
A second relational table is given: medicare_providers.tsv

8. Load the ~/Day4/datasets/medicare_providers.tsv file in R

```{r}
providers <- read_tsv("~/Day4/datasets/medicare_providers.tsv")
```


9. Use left_join() to merge the providers by name specialty with the claims dataframe

```{r}
joined_claims_bup <- left_join(buprenorphine_claims, providers, by = "provider_ID")
```


10. Which specialty prescribes buprenorphine the most.

```{r}
joined_claims_bup %>%
  count(provider_specialty) %>%
  arrange(desc(n))
```


Lastly, we will get the cost of buprenorphine per claim.


11. Using mutate(), add a new column dividing the values of total_drug_cost by total_claim_count. Call this column cost_per_claim

```{r}
joined_claims_bup %>%
  mutate(cost_per_claim = total_drug_cost/total_claim_count)
```


12. Using group_by() and summarize(), find the mean the cost_per_claim column per each state.

```{r}
joined_claims_bup %>%
  mutate(cost_per_claim = total_drug_cost/total_claim_count) %>%
  group_by(provider_state) %>%
  summarize(mean_per_state=mean(cost_per_claim))
```


13. Sort this output using arrange()

```{r}
joined_claims_bup %>%
  mutate(cost_per_claim = total_drug_cost/total_claim_count) %>%
  group_by(provider_state) %>%
  summarize(mean_per_state = mean(cost_per_claim)) %>%
  arrange(mean_per_state)
```


14. Lastly, use the output you just generated to print find out which state has the highest
  average cost_per_claim for buprenorphine

```{r}
joined_claims_bup %>%
  mutate(cost_per_claim = total_drug_cost/total_claim_count) %>%
  group_by(provider_state) %>%
  summarize(mean_per_state = mean(cost_per_claim)) %>%
  arrange(mean_per_state) %>%
  tail(1)
```

